Stateless Model Checking with Data-Race Preemption Points

### 2015-11-14 pldi-16/evaluation.tex
implement joshua's revisions pt1
Write eval for free-re-mallocs and totalverif

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{join\_test}}
\newcommand\bct{\texttt{bcast\_test}}
\newcommand\paraguay{\texttt{signal\_test}}
\newcommand\paradise{\texttt{sem\_test}}
\newcommand\rwldgr{\texttt{rwlock\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{sched\_test}}
\newcommand\waitsimple{\texttt{wait\_test}}
\newcommand\alarmsimul{\texttt{alarm\_test}}
% TODO: Add more test cases
We tested the Pintoses with 3 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented, so each test could only be run on a subset of the 78 submissions; see Table~\ref{tab:drbugs}.
}
The source code of all 9 tests is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
Finally, all tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
In total, the test suite comprises 629 unique state spaces,
% TODO: Add number of pintos bugs.
at least 106 of which contain bugs.

@@

\begin{table*}[t]
	\begin{center}
		%\small
	\begin{tabular}{r|c||c|c|c||c|c|c|c|c}
		% TODO: Make the "bugs" left half of this graph into a bar graph.
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Total} & {\bf Untested} & {\bf Malloc-} & {\bf Nondet.} & {\bf Nondet.} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR PPs} & {\bf DR PPs} & {\bf recycle DRs} & {\bf DR PPs} & {\bf DR bugs} \\
		\hline
		{\bct		} & 79	& 7	& 11	& 4	&	&	& 65	&	& 4	\\
		{\tej		} & 79	& 14 	& 25	& 11	&	&	& 471	&	& 6	\\
		{\mxtest	} & 79	& 1	& 10	& 9	&	&	& 7	&	& 1	\\
		{\paradise	} & 79	& 10	& 17	& 7	&	&	& 268	&	& 3	\\
		{\paraguay	} & 79	& 4	& 12	& 9	&	&	& 188	&	& 0	\\
		{\rwldgr	} & 79	& 26	& 30	& 3	&	&	& 157	&	& 3	\\
		\hline
		{\prisema	} & 59	&	&	&	&	&	&	&	&	\\
		{\waitsimple	} & 52	&	&	&	&	&	&	&	&	\\
		{\alarmsimul	} & 44	&	&	&	&	&	&	&	&	\\
		\hline
		{\bf Total}	& 629	&	&	&	&	&	&	&	&
	\end{tabular}
	\end{center}
	\caption{Summary of number and types of bugs found by each test case.
		SSS-MC is the control; QS is \quicksand.
		``DR-only'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}}).
	``Nondeterministic DRs'' counts among the DR-only bugs how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
		TODO: more explanation}
	\label{tab:drbugs}
\end{table*}

The left half of Table~\ref{tab:drbugs} breaks down the number and types of bugs found by each test case.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
even compared to the other test cases\footnote{
% "Aren't all lock impls assembly?" "Yes, but this one was ALL assembly."
	The one bug SSS-MC found was in a fully-assembly lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it suffices for many applications of MC to assume correctly-implemented locks,
we consider this strong evidence that any verification of low-level synchronization code must incorporate data-race PPs.

@@

We also counted the number of malloc-recycle false positives that \landslide~suppressed.

TODO

{\bf Full verification.}
TODO

@@

% TODO: Measure how much subset overlap there is.
{\bf Avoiding redundant work.}
When we extend a small state space with more PPs, the new state space is guaranteed to test a superset of thread interleavings compared to the old one.
Any interleaving which does not preempt threads on any of the new PPs will be repeated work.
Because we prioritize completing small state spaces before extending them with more PPs,
the superset state spaces we run later will always repeat each branch of their already-completed subsets.
%
We measured the proportion of repeated work among completed state spaces across our test suite;
on average, {\bf \large 999\%} of the interleavings in each test were repeated, with some tests as high as {\bf \large 9999\%}.

@@

% Future work: Add parallel DPOR so you can fill your spare CPUs when there are fewer than the max number of jobs.

>>>>

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{join\_test}}
\newcommand\bct{\texttt{bcast\_test}}
\newcommand\paraguay{\texttt{signal\_test}}
\newcommand\paradise{\texttt{sem\_test}}
\newcommand\rwldgr{\texttt{rwlock\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness,
\tej, a test of thread lifecycle,
\bct~and \paraguay~for cvars,
\paradise~for semaphores,
and \rwldgr~for r/w locks.
For \mxtest, \paradise, and \paraguay, we used {\tt without\_function} to blacklist {\tt thr\_create}, {\tt thr\_exit}, and {\tt thr\_join},
and for \mxtest~we enabled \landslide's mutex-testing option
(see \sect{\ref{sec:landslide}}).
\newcommand\prisema{\texttt{sched\_test}}
\newcommand\waitsimple{\texttt{wait\_test}}
\newcommand\alarmsimul{\texttt{alarm\_test}}
% TODO: Add more test cases
We tested the Pintoses with 3 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm,
\alarmsimul, for the timer sleep routine,
and \waitsimple, for process lifecycle system calls\footnote{
	Some of the Pintoses were partially implemented, so each test could only be run on a subset of the 78 submissions; see Table~\ref{tab:drbugs}.
}
The source code of all 9 tests is available at
{\em [submitted as supplementary material]}.
For all tests, we used {\tt without\_function} to blacklist PPs on the {\tt malloc} lock.
Finally, all tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
In total, the test suite comprises 629 unique state spaces,
at least 165 of which contain bugs.

@@

\begin{table*}[t]
	\begin{center}
		%\small
	\begin{tabular}{r|c||c|c|c||c|c|c|c|c}
		% TODO: Make the "bugs" left half of this graph into a bar graph.
		& {\bf Num.} & {\bf SSS-MC} & {\bf QS} & {\bf DR-only} & {\bf Total} & {\bf Untested} & {\bf Malloc-} & {\bf Nondet.} & {\bf Nondet.} \\
		{\bf Test name} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR PPs} & {\bf DR PPs} & {\bf recycle DRs} & {\bf DR PPs} & {\bf DR bugs} \\
		\hline
		{\bct		} & 79	& 7	& 11	& 4	&	&	& 65	&	& 4	\\
		{\tej		} & 79	& 14 	& 25	& 11	&	&	& 471	&	& 6	\\
		{\mxtest	} & 79	& 1	& 10	& 9	&	&	& 7	&	& 1	\\
		{\paradise	} & 79	& 10	& 17	& 7	&	&	& 268	&	& 3	\\
		{\paraguay	} & 79	& 4	& 12	& 9	&	&	& 188	&	& 0	\\
		{\rwldgr	} & 79	& 26	& 30	& 3	&	&	& 157	&	& 3	\\
		\hline
		{\prisema	} & 59	&	&	&	&	&	& 0	&	&	\\
		{\alarmsimul	} & 44	&	&	&	&	&	& 37	&	&	\\
		{\waitsimple	} & 52	&	&	&	&	&	& 57	&	&	\\
		\hline
		{\bf Total}	& 629	&	&	&	&	&	& 1250	&	&
	\end{tabular}
	\end{center}
	\caption{Summary of number and types of bugs found by each test case.
		SSS-MC is the control; QS is \quicksand.
		``DR-only'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}}).
	``Nondeterministic DRs'' counts among the DR-only bugs how many candidates required MC integration to identify (\sect{\ref{sec:eval-falseneg}}).
		TODO: more explanation}
	\label{tab:drbugs}
\end{table*}

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{bargraph.pdf}
	\caption{{\bf TODO:} Is this graph productive to include in the paper? Or should I just leave the numbers in table format (the left half of Table~\ref{tab:drbugs})?}
	\label{fig:bargraph}
\end{figure}

%Figure~\ref{fig:bargraph}
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test case.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
even compared to the other test cases\footnote{
% "Aren't all lock impls assembly?" "Yes, but this one was ALL assembly."
	The one bug SSS-MC found was in a fully-assembly lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it suffices for many applications of MC to assume correctly-implemented locks,
we consider this strong evidence that any verification of low-level synchronization code must incorporate data-race PPs.

@@

Beyond finding new bugs with data-race PPs, we evaluated \quicksand's performance in classifying data-race candidates in two ways.

{\bf Suppressing ``malloc-recycle'' false positives.}
In \sect{\ref{sec:recycle}} we showed the soundness of suppressing data race reports between two heap accesses when the surrounding memory was re-allocated in between.
In Table~\ref{tab:drbugs}, the column ``Malloc-recycle DRs'' shows the total number of such data-race candidates for each test case.
In total, 1250 data-races fit the malloc-recycle pattern across all tests,
only 69 of which were observed to avoid the re-allocation in an alternate interleaving.
Our proof in \sect{\ref{sec:recycle}} guarantees the safety of pruning all 1181 other state spaces.

Among the 69 true data-races which initially fit the malloc-recycle pattern,
none of them exposed a new bug when used as a PP.
This shows that for other data-race tools,
suppressing malloc-recycle candidates may be a productive heuristic,
even if unsound without Iterative Deepening.
However, \quicksand~was able to correctly identify 69 violations of the heuristic among 30 tests,
and fall back to classifying them with DPOR,
which we consider a much stronger verification.

{\bf Full verification.}
For 153 of our 629 tests, \quicksand~was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}}.
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of \quicksand's total verifications before a given elapsed CPU-time.
Among these, 36 tests contained no data-race candidates whatsoever, 
so the same verification could be provided by SSS-MC.
We plot these for comparison,
according to the time taken by that test's SSS-MC control experiment,.
When a test can be totally verified without any data-race PPs,
\quicksand~will be slower than SSS-MC due to redundant work (\sect{\ref{sec:future}}).
Of course, using data-race PPs increases our capacity for total verification by 4.25x.

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{totalverifs.pdf}
	\caption{Cumulative distribution of tests \quicksand~fully verified (\sect{\ref{sec:totalverif}}).
	Some tests had no data-race candidates, and hence could also be verified by SSS-MC.}
	\label{fig:totalverif}
\end{figure}

@@

% TODO CAMREADY: Measure how much subset overlap there is.
{\bf Avoiding redundant work.}
When we extend a small state space with more PPs, the new state space is guaranteed to test a superset of thread interleavings compared to the old one.
Any interleaving which does not preempt threads on any of the new PPs will be repeated work.
Because we prioritize completing small state spaces before extending them with more PPs,
the superset state spaces we run later will repeat each branch of their already-completed subsets.
%
%We measured the proportion of repeated work among completed state spaces across our test suite;
%on average, {\bf \large 999\%} of the interleavings in each test were repeated, with some tests as high as {\bf \large 9999\%}.

@@

Likewise, for test cases where full verification is not computationally feasible,
the state spaces associated with many data-race candidates will time out.
We cannot guarantee that those races are false positives or benign, even though no bug was found.
% FIXME: You need a separate table counting only the state spaces WITHOUT BUGES.
In the ``Untested DR PPs'' column of Table~\ref{tab:drbugs}, we count the number of such unverified candidates from each test case.
For a more formal treatment of these cases, we refer the reader to the {\em k-witness harmless} metric introduced by \cite{portend},
which could be combined with Iterative Deepening in future work.
% Future work: Add parallel DPOR so you can fill your spare CPUs when there are fewer than the max number of jobs.
